---
title: "p8105_hw3_zl3386"
author: "Ziqiu Liu"
date: "2023-10-08"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Problem 1

This problem uses the Instacart data.
```{r}
library(p8105.datasets)
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

Write a short description of the dataset, noting the __size__ and __structure__ of the data, describing some __key variables__, and giving illstrative __examples__ of observations.

* The data has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row represents a product (with id `product_id`) from an instacart order (with id `order_id`). Other variables include the order in which the item was added to cart (`add_to_cart_order`), whether or not the item has been ordered by the user before (`reordered`), customer identifier (`user_id`), etc. A product is identified by its `product_name` (e.g. Bulgarian Yogurt, Cucumber Kirby), its `aisle` name (e.g. yogurt, fresh vegetables) and id (`aisle_id`), and its `department` (e.g. dairy eggs, produce).

* Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

* Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

* Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, including the number of times each item is ordered.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

* Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

This problem uses the BRFSS data.

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning:
```{r}
brfss_df = 
  brfss_smart2010 |>
  # format the data to use appropriate variable names
  janitor::clean_names() |> 
  rename(state = locationabbr, location = locationdesc) |>
  # focus on the "Overall Health” topic
  filter(topic == "Overall Health") |>
  # include only responses from “Excellent” to “Poor”
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor" )) |>
  # organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_df |>
  filter(year == 2002) |>
  group_by(state) |>
  count(location) |> 
  count(state) |>
  filter(n >= 7)

brfss_df |>
  filter(year == 2010) |>
  group_by(state) |>
  count(location) |> 
  count(state) |>
  filter(n >= 7)
```

* In 2002, there were 6 states (CT, FL, MA, NC, NJ, PA) that were observed at 7 or more locations; while in 2010, this number had increased to 14.

Construct a dataset that is limited to `Excellent` responses, and contains year, state, and a variable that averages the `data_value` across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years

```{r}
excellent_df =
  filter(brfss_df, response == "Excellent") |>
  select(year, state, data_value) |>
  group_by(state, year) |>
  mutate(mean_value = mean(data_value)) |>
  select(-data_value) |>
  unique() |>
  arrange(state, year)

excellent_df |>
  ggplot(aes(x = year, y = mean_value, color = state)) +
  labs(title = "Average `data_value` for `Excellent` response within states over time") +
  geom_line(aes(group = state))
  
```

* In general, the proportion of `Excellent` responses in most states has not shown a significant increasing or decreasing trend over the years.

Make a two-panel plot showing, for the years 2006, and 2010, distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |>
  filter(year %in% c(2006, 2010), state == "NY") |>
  select(year, location, response, data_value) |>
  #ggplot(aes(x = response, y = data_value, color = location)) +
  #geom_line(aes(group = location)) +
  #facet_grid(. ~ year)
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(title = "Distribution of `data_value` for responses among locations in NY State") +
  facet_grid(. ~ year)
```

* Compared to 2006, there has been a significant increase in the proportion of `Very good` responses among locations in NY State in 2010.

## Problem 3

This problem uses accelerometer data collected on 250 participants in the NHANES study.

Load, tidy, merge, and otherwise organize the data sets.

Final dataset includes all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data.

```{r}
nhanes_covar =
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  rename(id = seqn) |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    ),
    education = case_match(
      education,
      1 ~ "< high school",
      2 ~ "high school",
      3 ~ "> high school"
    )
  ) |>
  mutate(education = factor(education, levels = c("< high school", "high school", "> high school"))) |>
  filter(age >= 21) |>
  na.omit()
  
nhanes_accel =
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  rename(id = seqn)

nhanes_df =
  left_join(nhanes_covar, nhanes_accel, join_by(id))
```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category.

```{r}
nhanes_df |>
  group_by(education) |>
  count(sex) |>
  knitr::kable()

nhanes_df |>
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  labs(title = "Age distributions for men and women in each education category") +
  facet_grid(. ~ education)
```

* From the table, we can see that the participants' sex ratio are mostly balanced in each education category, although in the high school equivalent education category there are more male participants than female (35:23). The more than high school education category contains most participants (56 male, 59 female).

* From the boxplot, we can see that participants with more than high school education is significantly younger than the other two education categories. In the high school equivalent education category, female participants are younger than than male participants, while in the other two categories the age distributions are similar between sexes.

Using the tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis).

```{r}
nhanes_df |>
  mutate(total_act = rowSums(nhanes_df[,6:1445])) |>
  relocate(id, sex, age, bmi, education, total_act) |>
  ggplot(aes(x = age, y = total_act, color = sex)) +
  geom_point() +
  geom_smooth(aes(group = sex), se = FALSE) +
  labs(title = "Total activities - Age") +
  facet_grid(. ~ education)
```

* In general, the amount of activity over a day goes down with the increase of age, and this trend is most significant in the less than high school education category. In high school equivalent and more than high school education categories, female participants tend to have more activity than male, while in less than high school category the difference is not significant.

Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.

```{r}
nhanes_df |>
  pivot_longer(
    cols = starts_with("min"),
    names_to = "time",
    values_to = "activity"
  ) |>
  mutate(time = as.numeric(str_remove(time, "min"))) |>
  ggplot(aes(x = time, y = activity, color = sex)) +
  geom_point(alpha = 0.1) +
  geom_smooth(aes(group = sex), se = FALSE, alpha = 1) +
  labs(
    title = "24-hour activity time courses for each education level",
    x = "time (h)"
    ) +
  scale_x_continuous(
    breaks = seq(0, 1440, by = 180),
    labels = as.character(seq(0, 24, by = 3))
    ) +
  facet_grid(. ~ education) +
  theme(legend.position = "bottom") 
```

* From the plot, we can see that in general, the activity level increases from a low level in the morning (6 ~ 12 h), remains stable in the afternoon and evening (12 ~ 21), and goes back to low level at night (21 ~ 6 h next day).