---
title: "p8105_hw3_zl3386"
author: "Ziqiu Liu"
date: "2023-10-08"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

This problem uses the Instacart data.
```{r}
library(p8105.datasets)
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

Write a short description of the dataset, noting the __size__ and __structure__ of the data, describing some __key variables__, and giving illstrative __examples__ of observations.

* The data has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row represents a product (with id `product_id`) from an instacart order (with id `order_id`). Other variables include the order in which the item was added to cart (`add_to_cart_order`), whether or not the item has been ordered by the user before (`reordered`), customer identifier (`user_id`), etc. A product is identified by its `product_name` (e.g. Bulgarian Yogurt, Cucumber Kirby), its `aisle` name (e.g. yogurt, fresh vegetables) and id (`aisle_id`), and its `department` (e.g. dairy eggs, produce).

* Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

* Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

* Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, including the number of times each item is ordered.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

* Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```



